How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-29 tse-paper/content.tex
Add todos for Caro and myself
spellchecker fixes

In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that these automatic software
builds create.
As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over
time~\cite{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

@@

% turned into an itemize again cause the 'sentence' around it was
% rather broken and itemish anyways
To fill this gap, we adapted and implemented three novel, prototypical
implementations to
automatically analyze build logs (step 2).
The ideas behind the implementations are based on promising approaches
adopted from other research fields and existing build log analysis
techniques.
In this article, we present three techniques that analyze a build log
by retrieving a substring (chunk) that describes a
specific information:
\begin{itemize}
	\item \emph{Program Synthesis by Example (PBE)}, a technique
	able to
	automatically synthesize regular expressions,
	\item a technique based on information
	retrieval, using \emph{Common Text Similarity (CTS)}
	\item inspired by the manual adhoc-search for key parts in the
	log, a
	technique employing \emph{Keyword Search (KWS)}
\end{itemize}
% a technique 1) able to automatically synthesize regular expressions,
% called \emph{Program Synthesis by Example (PBE)}, a technique
% 2) based on information
% retrieval, using \emph{Common Text Similarity (CTS)}	and 3)
% inspired by the manual adhoc-search for key parts in the log, a
% technique employing \emph{Keyword Search (KWS)}.
These three approaches---synthesized regular expressions,
text similarity,
keyword search---have
different strengths and weaknesses.
As a consequence,
in the second research question we ask:

@@

\subsubsection{Which kinds of build logs are subject to
  analysis? (RQ 1.3)}
\Cref{fig:litsur:log_producer} shows the distribution of supported
build tools.
Of the articles we surveyed, 30\% analyzed
Mave~\cite{maven2019website},
16\% analyzed Gradle~\cite{gradle2020website},
and 10\% analyzed Ant build logs~\cite{ant2020website}.

@@

% TODO Caro: adapt as discussed
% TODO: I think the axes labels are also unncessary for this graph.
%Our other graphs in the lit survey do not have them, either.
\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

@@

% TODO merge nicely
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.

@@

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
The user provides these examples and keywords, which are easy to adapt
whenever the user adds new examples.
A user---a developer or a researcher who wants to analyze
build logs---provides these examples and keywords, which can simply be

@@

We implement three chunk retrieval techniques based on approaches
previously described by several articles:
PBE mimics regular expressions, while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad-hoc approach of
searching for specific strings or keywords within KWS\@.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
% TODO Caro: Wouldn't this place perhaps be better to contain
%the pargraph about why chunk log techniques are more general
% than other approaches?

\label{sec:crt-characteristics}
In this section, we formalize our use of chunk retrieval techniques.
By this word, we refer to techniques that automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.
% and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
% TODO TF-IDF undefined
algorithm weighs the vectors using TF-IDF, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsection{Keyword Search (KWS)}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
% TODO: unclear
We associate each example with keywords which appear in the targeted
chunk or close to it.

% TODO: unclear.
or do you mean: same number == all searched for ?
KWS then searches for those keywords which are tied to the greatest
number of examples in the training set.
If several keywords are associated with the same number of training
examples, and no other keywords are associated with more training
examples, KWS searches for all of these keywords.

@@

\begin{figure}[tbp]
  \centering
\begin{subfigure}[tbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

@@

\begin{subfigure}[tbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \lstinputlisting[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]{listings/chunk2.txt}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error.}
  \label{lst:chunk-example-3}
\end{subfigure}

\caption{Examplary log excerpts.}
\label{lst:logexamples}
\end{figure}

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition,
RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[tbp]
  \centering
  \lstinputlisting[breaklines=true]{listings/prose.txt}
  \caption{Simplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example}.}
  \label{lst:prose-program-simplified}
\end{figure}

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3}
to the training set (which has a
different structural representation),
PBE is not able to synthesize a program anymore and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
separating log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
% TODO Moritz: is this fitting here?
% or too much 'foreshadowing' too far down in the hierachy
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

\end{subfigure}\hspace{\fill}
\begin{subfigure}[tb]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{structural category count
		in training and test set.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall, and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS) compared by \ldots}
% TODO: where are the bean plots?
\label{fig:results-CTS}
\end{figure*}

% TODO
% You might even say Figure 12-15 show two
% plots in the short text under the section heading.
% It just makes the
% text flow so much better.

@@

% we can't just have a paragraph in a scientific paper start with
% Caveat, I think ;-) Also, we should say why we THINK this is not
%the case.
% Otherwise we weaken our work unduly
% Of course we can! xD
% Agree with the rest.
% It's in TtV, quite close actually
% so I'll comment this out
% also #moreselling :D
% \textbf{Caveat!}
% This is a preliminary hypothesis based on the results
% from our comparison study.
% The recommendations could therefore be
% influenced by idiosyncrasies of our specific implementation of the
% chunk retrieval techniques as well as the \emph{LogChunks}
% data set.

>>>>

\IEEEPARstart{I}{n}
the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that these automatic software
builds create.
As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over
time~\cite{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

@@

% turned into an itemize again cause the 'sentence' around it was
% rather broken and item-ish anyways
To fill this gap, we adapted and implemented three novel, prototypical
implementations to
automatically analyze build logs (step 2).
The ideas behind the implementations are based on promising approaches
adopted from other research fields and existing build log analysis
techniques.
In this article, we present three techniques that analyze a build log
by retrieving a substring (chunk) that describes a
specific information:
\begin{itemize}
	\item \emph{Program Synthesis by Example (PBE)}, a technique
	able to
	automatically synthesize regular expressions,
	\item a technique based on information
	retrieval, using \emph{Common Text Similarity (CTS)}
	\item inspired by the manual ad hoc search for key parts in the
	log, a
	technique employing \emph{Keyword Search (KWS)}
\end{itemize}
% a technique 1) able to automatically synthesize regular expressions,
% called \emph{Program Synthesis by Example (PBE)}, a technique
% 2) based on information
% retrieval, using \emph{Common Text Similarity (CTS)}	and 3)
% inspired by the manual ad hoc search for key parts in the log, a
% technique employing \emph{Keyword Search (KWS)}.
These three approaches---synthesized regular expressions,
text similarity,
keyword search---have
different strengths and weaknesses.
As a consequence,
in the second research question we ask:

@@

\subsubsection{Which kinds of build logs are subject to
  analysis? (RQ 1.3)}
\Cref{fig:litsur:log_producer} shows the distribution of supported
build tools.
Of the articles we surveyed, 30\% analyzed
Maven~\cite{maven2019website},
16\% analyzed Gradle~\cite{gradle2020website},
and 10\% analyzed Ant build logs~\cite{ant2020website}.

@@

% TODO Caro: adapt as discussed
% TODO Caro: I think the axes labels are also unnecessary for this graph.
%Our other graphs in the lit survey do not have them, either.
\begin{figure}[tbhp]
		\centering
		\includegraphics[width=\columnwidth, trim={1.1cm 0.4cm
		1.5cm 0.5cm},
		clip]{img/lit-sur/log_producer_annotated.pdf}
		\caption{Frequency of supported log producers.}
		\label{fig:litsur:log_producer}
\end{figure}

@@

% TODO Caro: merge nicely
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.

@@

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
% TODO Caro: merge!
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
The user provides these examples and keywords, which are easy to adapt
whenever the user adds new examples.
A user---a developer or a researcher who wants to analyze
build logs---provides these examples and keywords, which can simply be

@@

We implement three chunk retrieval techniques based on approaches
previously described by several articles:
PBE mimics regular expressions, while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad hoc approach of
searching for specific strings or keywords within KWS\@.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
% TODO Caro: Wouldn't this place perhaps be better to contain
%the paragraph about why chunk log techniques are more general
% than other approaches?

\label{sec:crt-characteristics}
In this section, we formalize our use of chunk retrieval techniques.
By this word, we refer to techniques that automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.
% and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad hoc extracting just one specific piece of information per
configuration.

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
% TODO Caro: TF-IDF undefined
algorithm weighs the vectors using TF-IDF, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsection{Keyword Search (KWS)}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
% TODO Caro: unclear
We associate each example with keywords which appear in the targeted
chunk or close to it.

% TODO Caro: unclear.
or do you mean: same number == all searched for ?
KWS then searches for those keywords which are tied to the greatest
number of examples in the training set.
If several keywords are associated with the same number of training
examples, and no other keywords are associated with more training
examples, KWS searches for all of these keywords.

@@

% TODO Moritz: reference these in text properly

\lstset{escapeinside=//}
\begin{figure}[tbp]
  \centering
\begin{subfigure}[tbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
FAILURE: Build failed with an exception.

@@

\begin{subfigure}[tbp]{\columnwidth}
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \lstinputlisting[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]{listings/chunk2.txt}
  \vspace{-\baselineskip}
  % TODO Caro: Adapt frame= option to close frame properly
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error.}
  \label{lst:chunk-example-3}
\end{subfigure}

\caption{Exemplary log excerpts.}
\label{lst:logexamples}
\end{figure}

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition,
RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}}
\begin{figure}[tbp]
  \centering
  \lstinputlisting[breaklines=true]{listings/prose.txt}
  \caption{Simplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example}.}
  \label{lst:prose-program-simplified}
\end{figure}

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3}
to the training set (which has a
different structural representation),
PBE is not able to synthesize a program anymore and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
separating log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
% TODO Moritz: is this fitting here?
% or too much 'foreshadowing' too far down in the hierarchy
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

\end{subfigure}\hspace{\fill}
\begin{subfigure}[tb]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{structural category count
		in training and test set.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall, and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS) compared by \ldots}
% TODO Caro: check: where are the bean plots?
\label{fig:results-CTS}
\end{figure*}

% TODO Caro: check
% You might even say Figure 12-15 show two
% plots in the short text under the section heading.
% It just makes the
% text flow so much better.

@@

% we can't just have a paragraph in a scientific paper start with
% Caveat, I think ;-) Also, we should say why we THINK this is not
%the case.
% Otherwise we weaken our work unduly
% Of course we can! xD
% Agree with the rest.
% It's in TtV, quite close actually
% so I'll comment this out
% also #more selling :D
% \textbf{Caveat!}
% This is a preliminary hypothesis based on the results
% from our comparison study.
% The recommendations could therefore be
% influenced by idiosyncrasies of our specific implementation of the
% chunk retrieval techniques as well as the \emph{LogChunks}
% data set.
